Storm Surges

### 2014-04-23 RoughWork/manuscript.tex
Added description of NEMO, SSH equation, eddy viscosities.

\documentclass[pdftex,12pt]{article}
\usepackage{fullpage}
\usepackage{graphicx}
\usepackage{natbib}
\usepackage{amsmath}

@@

\section{Model Configuration}\label{sec:config}
%A section about how we configured the model for SoG 
\subsection{Model domain}
The modelled domain extends from the Strait of Juan de Fuca to Puget Sound to Johnstone Strait as shown in Figure \ref{fig:domain}. Bathymetry from the Cascadia physiography dataset \citep{haugerud1999digital} was smoothed to limit the difference in depth across grid cells. For model stability, additional smoothing at the Strait of Juan de Fuca western boundary was imposed to achieve constant depth across the first ten grid cells.

@@

\subsection{Boundary conditions}
%lateral and bottom friction, open boundaries, eta surface

@@

The monthly flow from each river was input as a point source in the three grid points closest to the surface at the model point closest to the mouth of each river. Incoming water was assumed to be fresh and at the temperature of the surface. A total of 150 rivers were parameterised by this method.

@@

\subsection{Initial conditions}
Initial conditions for temperature and salinity were taken from a CTD cast in the middle Strait of Georgia taken in Sept 2002 \citep{pawlowiczetal2007}.  Conditions were initially uniform horizontally.  Velocity was initialized at zero.

>>>>

\documentclass[pdftex,10pt]{article}
\usepackage{fullpage}
\usepackage{graphicx}
\usepackage{natbib}
\usepackage{amsmath}

@@

\section{Model Configuration}\label{sec:config}
%A section about how we configured the model for SoG 
We have used the Nucleus for European Modelling of the Ocean (NEMO) framework in its regional configuration to develop an ocean model for the Strait of Georgia and Salish Sea. NEMO is a highly modularized tool used for studying ocean physics, ocean-ice interactions, and the biogeochemical properties of the ocean. NEMO's ocean core solves the three-dimensional hydrostatic equations of motion for an incompressible fluid under the Boussinesq approximation on a structured computational grid. Although not used in the present work, NEMO's options for grid nesting and biogeochemical coupling make it a useful tool for studying the complex physics and biogeochemical interactions within the Strait of Georgia. This work focuses on validating the physical set up of the Salish Sea model, in particular, determining appropriate forcing and boundary conditions for accurate reproduction of tidal amplitudes and phases as well as storm surge elevations. Future work will include biogeochemical coupling and data assimilation.

\subsection{Model domain}
The modelled domain extends from the Strait of Juan de Fuca to Puget Sound to Johnstone Strait as shown in Figure \ref{fig:domain}. Bathymetry from the Cascadia physiography dataset \citep{haugerud1999digital} was smoothed to limit the difference in depth across grid cells. For model stability, additional smoothing at the Strait of Juan de Fuca western boundary was imposed to achieve constant depth across the first ten grid cells. As depicted in Figure \ref{fig:domain}, the numerical grid is rotated $29^{\circ}$ counter-clockwise of North in order to maintain computational efficency since currents within the Strait of Georgia are mainly aligned with this rotated axis.

The curvilinear orthogonal numerical grid is divided into 400 by 900 by 40 grid cells, which results in an almost uniform horizontal resolution with grid spacing approximately 400 m by 500 m. The 40 vertical $z$-levels are stretched gradually in order to achieve higher resolution in the surface layer, with approximately 1 m vertical grid spacing down to about 20 m in depth. At the bottom boundary, partial $z$-levels are utilized in order to limit large changes in bathymetry across grid cells \citep{madec2008nemo}.

@@

\subsection{Boundary conditions and subgrid scale processes}
%lateral and bottom friction, open boundaries, eta surface
The model includes two open boundaries that connect to the Pacific Ocean, the western boundary of the Strait of Juan de Fuca as well as Johnstone Strait at the north, both of which are forced with eight tidal constituents, temperature and salinity climatologies, and the sea surface height anomaly. The details of these forcing conditions are described below. At coastal boundaries, the partial slip boundary condition, an approximation of the no slip boundary condition, is used. The partial slip boundary condition allows one to include the frictional effects of lateral boundaries without the restricitve resolution required to represent the lateral boundary layer under no slip conditions. A lateral eddy viscosity of 20 m $^2$ s $^{-1}$ parameterizes horizontal friction and the lateral eddy diffusivity is 20.5 m $^2$ s $^{-1}$ . At the ocean surface, meterological conditions and fresh water input from rivers are included and described in more detail below. Bottom friction is represented by a quadratic law for the bottom momentum flux with drag coefficient $C_D = 5\times 10^{-3}$. Vertical turbulence and mixing is calculated through the $k-\epsilon$ configuration of the generic length scale (GLS) turbulence closure \citep{umlauf2003generic} with background vertical eddy viscosity and diffusivity set to $1\times10^{-4}$ and $1\times10^{-5}$ m$^2$ s$^{-1}$ respectively. Details on the NEMO implementation of the partial slip lateral boundary condition, quadratic bottom friction law, and GLS turbulence closure scheme are provided by \citep{madec2008nemo}.

In addition to the equations of motion, a prognostic equation for the sea surface height is solved at each time step. The inclusion the sea surface height equation requires a fairly restrictive time step due to the presence of high speed surface gravity waves. As such, the split-explicit time stepping algorithm is employed, where the free surface and barotropic equations are solved with a smaller time step than that used for the other variables. The model time step and barotropic time step are 10 s and 2 s respectively.

@@

The monthly flow from each river was input as a point source in the three grid points closest to the surface at the model point closest to the mouth of each river. Incoming water was assumed to be fresh and at surface temperatyre. A total of 150 rivers were parameterised by this method.

@@

\subsection{Initial conditions}
Initial conditions for temperature and salinity were taken from a CTD cast in the middle Strait of Georgia taken in Sept 2002 \citep{pawlowiczetal2007}.  Conditions were initially uniform horizontally.  Velocity was initialized at zero. 
%Merge with below?
%Perhaps image of spun-up stratifiction along Thalweg? Here or in model evaluation
